```{r}
nftools::nf_library(c(
  'lubridate',
  'tidyverse',
  'data.table'
))

base <- '~/src/Sepsis-R01/Data_All_2023-12-21/'

dx <- fread(paste0(base, 'dx_all.csv'))
ed_keys <- fread(paste0(base, 'ed_keys.csv'))
enc_dx <- fread(paste0(base, 'enc_dx.csv'))
flowsheet <- fread(paste0(base, 'flowsheet.csv'))
hosp_adm <- fread(paste0(base, 'hosp_adm.csv'))
old_labs <- fread(paste0(base, 'labs.csv'))
pt <- fread(paste0(base, 'pt.csv')) %>%
  mutate(
    has_birthdate = ifelse(is.na(BirthDate), F, T)
  )
durationdim <- fread(paste0('~/src/Sepsis-R01/', 'durationdim.csv')) %>% 
  select(DurationKey, Years) %>%
  rename(DurationAge=Years)
ed <- fread(paste0(base, 'ed.csv')) %>%
  merge(durationdim, by.x = 'AgeKey', by.y='DurationKey', all.x=T)

# Sort by timestamp
flowsheet <- flowsheet %>%
  mutate(timestamp = ymd_hm(paste(DateKeyValue, TimeOfDayKeyValue))) %>%
  dplyr::filter(EncounterKey %in% ed_keys$EncounterKey)

ed <- ed %>%
  mutate(ed_timestamp = ymd_hm(paste(ArrivalDateKeyValue, ArrivalTimeOfDayKeyValue)))

# Load supporting files
base <- '~/src/Sepsis-R01/Data_All_2_2023-12-27/'
death <- fread(paste0(base, 'death.csv'))
#ed <- fread(paste0(base, 'ed.csv'))
enc_keys <- fread(paste0(base, 'enc_keys.csv'))
flowsheet_assess <- fread(paste0(base, 'flowsheet_assess.csv'))
flowsheet_resp <- fread(paste0(base, 'flowsheet_resp.csv'))
#hosp_adm <- fread(paste0(base, 'hosp_adm.csv'))
icu_reg <- fread(paste0(base, 'icu_reg.csv'))
insurance <- fread(paste0(base, 'insurance.csv'))
labs <- fread(paste0(base, 'labs.csv'))
med_ab <- fread(paste0(base, 'med_ab.csv'))
med_vp <- fread(paste0(base, 'med_vp.csv'))

#pt <- fread(paste0(base, 'pt.csv'))
pt_keys <- fread(paste0(base, 'pt_keys.csv'))

# Sort by collection time, not result time
labs <- labs %>%
  mutate(timestamp = ymd_hm(paste(CollectionDateKeyValue, CollectionTimeOfDayKeyValue)))
```

```{r WBC}
wbc <- labs %>%
  mutate(
    parsed_value = parse_number(Value),
    is_na_coercion = is.na(parsed_value)
  )
print(wbc %>% ungroup %>% filter(is_na_coercion) %>% count(Value))

wbc <- wbc %>%
  mutate(NumericValue = as.numeric(Value)) %>%
  filter(ComponentName== 'WBC Count' & !is.na(NumericValue)) %>%
  group_by(EncounterKey) %>%
  arrange(timestamp) %>%
  slice(1) %>%
  merge(ed %>% select(EncounterKey, ed_timestamp), by='EncounterKey', all.x=T) %>%
  mutate(
    TWO_HOURS = ifelse(difftime(timestamp, ed_timestamp, units = "hours") < 2, 1, 0),
    SIRS_WBC = ifelse(TWO_HOURS & (NumericValue > 12 | NumericValue < 4), 1, 0)
  ) %>%
  rename(wbc = NumericValue)

wbc %>% ungroup %>% count(SIRS_WBC)
```

```{r Temperature}
temp <- flowsheet %>%
  filter(FlowsheetRowDisplayName == 'Temp' & !is.na(NumericValue) & Accepted == 1) %>%
  mutate(temp = ifelse(NumericValue > 42, (NumericValue - 32) * 5/9, NumericValue)) %>%
  filter(temp >= 32 & temp <= 42) %>%
  group_by(EncounterKey) %>%
  arrange(timestamp) %>%
  slice(1) %>%
  merge(ed %>% select(EncounterKey, ed_timestamp), by='EncounterKey', all.x=T) %>%
  mutate(
#    TEMP_TWO_HOURS = ifelse(difftime(timestamp, ed_timestamp, units = "hours") < 2, 1, 0),
#    SIRS_TEMP = ifelse(TEMP_TWO_HOURS & (temp > 38 | temp < 36), 1, 0)
    SIRS_TEMP = ifelse(temp > 38 | temp < 36, 1, 0)
  )

temp %>% ungroup %>% count(SIRS_TEMP)
```


```{r}
df <- ed_keys %>%
  merge(wbc %>% select(PatientDurableKey, EncounterKey, wbc, TWO_HOURS, SIRS_WBC, timestamp), by='EncounterKey', all.x=T, suffixes=c('_ED', '_WBC')) %>%
  merge(temp %>% select(PatientDurableKey, EncounterKey, temp, SIRS_TEMP, timestamp), by='EncounterKey', all.x=T, suffixes=c('', '_TEMP')) %>%
  mutate(
    analyzed = ifelse(!is.na(wbc), T, ifelse(!is.na(temp), T, F)),
    summation = rowSums(select(., contains('SIRS')), na.rm = TRUE),
    include = ifelse(summation > 0, T, F)
    )

write.csv(df, file='inclusion.csv')
```

```{r Define Sepsis ICD Codes}
icd_codes <- c(dx %>% 
                 filter(grepl('sepsis|septic', icd_l3_name, ignore.case=T)) %>% 
                 pull(icd_value) %>% 
                 unique,
               dx %>% 
                 filter(grepl('sepsis|septic', icd_l4_name, ignore.case=T)) %>% 
                 pull(icd_value) %>% 
                 unique,
               dx %>% 
                 filter(grepl('sepsis|septic', icd_l5_name, ignore.case=T)) %>%
                 pull(icd_value) %>% 
                 unique,
               dx %>% 
                 filter(grepl('sepsis|septic', icd_l6_name, ignore.case=T)) %>%
                 pull(icd_value) %>% 
                 unique
) %>% unique %>% sort

shock_icd_codes <- c(
  dx %>% 
    filter(grepl('sepsis|septic', icd_l3_name, ignore.case=T) & 
             grepl('shock', icd_l3_name, ignore.case=T) & 
             !grepl('without', icd_l3_name, ignore.case=T)) %>% 
    pull(icd_value) %>% 
    unique,
  dx %>% 
    filter(grepl('sepsis|septic', icd_l4_name, ignore.case=T) & 
             grepl('shock', icd_l4_name, ignore.case=T) & 
             !grepl('without', icd_l4_name, ignore.case=T)) %>% 
    pull(icd_value) %>% 
    unique,
  dx %>% 
    filter(grepl('sepsis|septic', icd_l5_name, ignore.case=T) & 
             grepl('shock', icd_l5_name, ignore.case=T) & 
             !grepl('without', icd_l5_name, ignore.case=T)) %>% 
    pull(icd_value) %>% 
    unique,
  dx %>% 
    filter(grepl('sepsis|septic', icd_l6_name, ignore.case=T) & 
             grepl('shock', icd_l6_name, ignore.case=T) & 
             !grepl('without', icd_l6_name, ignore.case=T)) %>% 
    pull(icd_value) %>% 
    unique,
  dx %>% 
    filter(grepl('sepsis|septic', icd_l7_name, ignore.case=T) & 
             grepl('shock', icd_l7_name, ignore.case=T) & 
             !grepl('without', icd_l7_name, ignore.case=T)) %>% 
    pull(icd_value) %>% 
    unique
) %>% unique %>% sort
```


```{r}
dx_trim <- dx %>%
  select(
    'EncounterKey',
    'icd_value'
  ) %>%
  mutate(
    sepsis = ifelse(icd_value %in% icd_codes, 1, 0),
    septic_shock = ifelse(icd_value %in% shock_icd_codes, 1, 0)
  ) %>%
  filter(EncounterKey %in% ed_keys$EncounterKey) %>%
  group_by(EncounterKey) %>%
  summarize(
    sepsis = max(sepsis),
    septic_shock = max(septic_shock)
  ) %>%
  mutate(
    sepsis = ifelse(sepsis == 1, T, F),
    septic_shock = ifelse(septic_shock == 1, T, F),
  ) %>%
  ungroup

pt_trim <- pt %>%
  select(
    'BirthDate',
    'has_birthdate',
    'PatientDurableKey',
    'Status',
    'MaritalStatus',
    'Religion',
    'SmokingStatus',
    'PrimaryFinancialClass',
    'HighestLevelOfEducation',
    'SexAssignedAtBirth',
    'GenderIdentity',
    'SexualOrientation',
    'UCSFDerivedRaceEthnicity_X'
  ) %>%
  distinct(.keep_all = TRUE) %>%
  group_by(PatientDurableKey) %>%
  slice(1) %>%
  ungroup

tabledf <- df %>%
  merge(dx_trim, by='EncounterKey', all.x = T) %>%
  merge(ed %>% dplyr::select(EncounterKey, DurationAge), by='EncounterKey', all.x = T) %>% 
  merge(pt_trim, by='PatientDurableKey', all.x = T) %>%
  mutate(
    processed_birthdate = lubridate::ymd(BirthDate),
    age_days = as.numeric(difftime(lubridate::ymd(ArrivalDateKeyValue), processed_birthdate, units='days')),
    age = age_days/365.25,
    age_delta = age - DurationAge,
    SIRS_WBC = as.factor(SIRS_WBC),
    SIRS_TEMP = as.factor(SIRS_TEMP),
  ) %>%
  mutate_if(is.character, 
    list(~ ifelse(grepl("^\\*|^$|^Missing$", .), NA, .))
  )

columns <- c(
#    'Status',
#    'age',
    'analyzed',
    'SIRS_WBC',
    'SIRS_TEMP',
    'DurationAge',
#    'age_delta',
#    'has_birthdate',
    'SexAssignedAtBirth',
    'GenderIdentity',
    'MaritalStatus',
    'Religion',
    'SmokingStatus',
    'PrimaryFinancialClass',
    'SexualOrientation',
    'wbc',
    'temp',
    'sepsis',
    'septic_shock'
)

nftools::nf_table1(columns, 'include', tabledf, calculate_pvalue = F)

tabledf$year <- tabledf$ArrivalDateKeyValue%>% year
nftools::nf_table1(columns, 'year', tabledf, calculate_pvalue = F)
```
```{r Summary stats for R01 Writeup}
# Number of visits per year
tabledf %>% count(year) %>% pull(n) %>% summary
```
```{r Determine Flag Timestamp}
df_timestamp <- df %>%
  filter(include == T) %>%
  rename(timestamp_WBC=timestamp) %>%
  mutate(timestamp=ifelse(is.na(SIRS_TEMP) | SIRS_TEMP == 0,
                          timestamp_WBC,
                          ifelse(is.na(SIRS_WBC) | SIRS_WBC == 0,
                                 timestamp_TEMP,
                                 ifelse(timestamp_WBC < timestamp_TEMP,
                                        timestamp_WBC,
                                        timestamp_TEMP)))) %>%
  mutate(timestamp = as_datetime(ymd_hms("1970-01-01 00:00:00") + seconds(timestamp))) %>%
  select(contains('Key'), contains('timestamp'), contains('SIRS')) %>%
  mutate(
    begin_timestamp = timestamp + hours(48) - hours(4),
    end_timestamp = timestamp + hours(48) + hours(4),
  )

df_timestamp <- df_timestamp %>% select(
  'EncounterKey',
  'begin_timestamp', 
  'end_timestamp'
)

df <- df %>%
  merge(df_timestamp, by='EncounterKey', all.x=T)
```

```{r eSOFA Vasopressor}
esofa_pressors <- med_vp %>%
  filter(EncounterKey %in% ed_keys$EncounterKey) %>%
  mutate(timestamp = ymd_hm(paste(AdministrationDateKeyValue, AdministrationTimeOfDayKeyValue))) %>%
  merge(df_timestamp, by='EncounterKey', all.x=T) %>%
  filter(timestamp > begin_timestamp & timestamp < end_timestamp) %>%
  group_by(EncounterKey) %>%
  slice(1) %>%
  mutate(ESOFA_PRESSOR = 1) %>%
  select(EncounterKey, timestamp, ESOFA_PRESSOR)
```

> icu_reg %>% count(IntubationDays)
> resp %>% count(FlowsheetRowName) %>% as.data.table %>% arrange(-n)
                                FlowsheetRowName      n
 1:                                       R FIO2 692846
 2:                         R RESP PEEP/CPAP SET  96036 #Good
 3:                               R RT VENT MODE  92240 #Good
 4:                        R RT MODE/BREATH TYPE  91142 #Good but with DEID columns
 5:      R RT VENTILATOR RESPIRATORY RATE  TOTAL  90097
 6:                   RT MAP MEASURED_IP_CD_UCSF  89282
 7:                       R RT (ADULT) VENT TYPE  84673
 8:    R RT VENTILATOR MINUTE VOLUME MEASURED V2  70198
 9: R RT (ADULT) VENTILATOR TIDAL VOLUME EXHALED  58667
10:   R RT (ADULT) VENTILATOR I:E RATIO MEASURED  56879
11:                                R RESP RR SET  53608
12:              R OXYGEN SATURATION INDEX (OSI)  51269
13:                                     R FIO2 2   9678

```{r eSOFA Mechanical Ventilation}
esofa_mv <- resp %>%
  filter(EncounterKey %in% ed_keys$EncounterKey) %>%
  filter(FlowsheetRowName %in% c('R RESP PEEP/CPAP SET', 'R RT VENT MODE', 'R RT MODE/BREATH TYPE')) %>%
  mutate(timestamp = ymd_hm(paste(DateKeyValue, TimeOfDayKeyValue))) %>%
  merge(df_timestamp, by='EncounterKey', all.x=T) %>%
  filter(timestamp > begin_timestamp & timestamp < end_timestamp)
  arrange(timestamp) %>%
  group_by(EncounterKey) %>%
  slice(1) %>%
  mutate(ESOFA_MV = 1) %>%
  select(EncounterKey, timestamp, ESOFA_MV)
```

> labs %>% count(ComponentName)
                             ComponentName      n
 1:                      Bilirubin, Direct  12337
 2:                       Bilirubin, Total 139478
 3:                             Creatinine 352940
 4:                 Creatinine Whole Blood     35
 5:                      Creatinine, Serum     60
 6:                             Hemoglobin 334015
 7: Hemoglobin, Plasma, Alternative Method    182
 8:                        Lactate, plasma  15653
 9:                   Lactate, whole blood 117056
10:                                    PO2 161882
11:                            PO2, Venous     42
12:                 POCT, Total Hemoglobin    598
13:                         Platelet Count 336999
14:                       Total Hemoglobin      5
15:                              WBC Count 333037

```{r eSOFA Lactate}
esofa_lactate <- labs %>%
  filter(EncounterKey %in% ed_keys$EncounterKey) %>%
  filter(ComponentName %in% c('Lactate, plasma')) %>%
  mutate(timestamp = ymd_hm(paste(CollectionDateKeyValue, CollectionTimeOfDayKeyValue))) %>%
  merge(df_timestamp, by='EncounterKey', all.x=T) %>%
  filter(timestamp > begin_timestamp & timestamp < end_timestamp) %>%
  arrange(timestamp) %>%
  filter(Value >= 2) %>%
  mutate(ESOFA_LACTATE = 1) %>%
  select(EncounterKey, timestamp, ESOFA_LACTATE)
```

```{r Doubling of Creatinine}
baseline_creatinine <- labs %>%
  filter(EncounterKey %in% ed_keys$EncounterKey & ComponentName %in% c('Creatinine', 'Creatinine, Serum')) %>%
  mutate(
    NumericValue = as.numeric(Value),
    timestamp = ymd_hm(paste(CollectionDateKeyValue, CollectionTimeOfDayKeyValue))
  ) %>%
  arrange(timestamp) %>%
  group_by(EncounterKey) %>%
  slice(1) %>%
  select(EncounterKey, NumericValue) %>%
  rename(baseline_creatinine=NumericValue) %>%
  ungroup

esofa_creatinine <- labs %>%
  filter(EncounterKey %in% ed_keys$EncounterKey & ComponentName %in% c('Creatinine', 'Creatinine, Serum')) %>%
  mutate(
    NumericValue = as.numeric(Value),
    timestamp = ymd_hm(paste(CollectionDateKeyValue, CollectionTimeOfDayKeyValue))
  ) %>%
  merge(baseline_creatinine, by='EncounterKey', all.x=T) %>%
  merge(df_timestamp, by='EncounterKey', all.x=T) %>%
  arrange(timestamp) %>%
  filter(timestamp > begin_timestamp & timestamp < end_timestamp & NumericValue >= baseline_creatinine * 2) %>%
  mutate(ESOFA_CREATININE= 1) %>%
  select(EncounterKey, timestamp, ESOFA_CREATININE)
```

```{r Doubling of Bilirubin to ≥2 mg/dL}
baseline_bilirubin <- labs %>%
  filter(EncounterKey %in% ed_keys$EncounterKey & ComponentName %in% c('Bilirubin, Total')) %>%
  mutate(
    NumericValue = as.numeric(Value),
    timestamp = ymd_hm(paste(CollectionDateKeyValue, CollectionTimeOfDayKeyValue))
  ) %>%
  arrange(timestamp) %>%
  group_by(EncounterKey) %>%
  slice(1) %>%
  select(EncounterKey, NumericValue) %>%
  rename(baseline_bilirubin=NumericValue) %>%
  ungroup

esofa_bilirubin <- labs %>%
  filter(EncounterKey %in% ed_keys$EncounterKey & ComponentName %in% c('Bilirubin, Total')) %>%
  mutate(
    NumericValue = as.numeric(Value),
    timestamp = ymd_hm(paste(CollectionDateKeyValue, CollectionTimeOfDayKeyValue))
  ) %>%
  merge(baseline_bilirubin, by='EncounterKey', all.x=T) %>%
  merge(df_timestamp, by='EncounterKey', all.x=T) %>%
  arrange(timestamp) %>%
  filter(timestamp > begin_timestamp & timestamp < end_timestamp & NumericValue >= baseline_bilirubin * 2 & NumericValue >= 2) %>%
  mutate(ESOFA_BILIRUBIN = 1) %>%
  select(EncounterKey, timestamp, ESOFA_BILIRUBIN)
```

```{r ≥50% Decrease in Platelet Count to less than 100 cells/microL}
baseline_platelet <- labs %>%
  filter(EncounterKey %in% ed_keys$EncounterKey & ComponentName %in% c('Platelet Count')) %>%
  mutate(
    NumericValue = as.numeric(Value),
    timestamp = ymd_hm(paste(CollectionDateKeyValue, CollectionTimeOfDayKeyValue))
  ) %>%
  arrange(timestamp) %>%
  group_by(EncounterKey) %>%
  slice(1) %>%
  select(EncounterKey, NumericValue) %>%
  filter(NumericValue >= 100) %>% # Filter out baseline values < 100
  rename(baseline_platelet=NumericValue) %>%
  ungroup

esofa_platelet <- labs %>%
  filter(EncounterKey %in% ed_keys$EncounterKey & ComponentName %in% c('Platelet Count')) %>%
  mutate(
    NumericValue = as.numeric(Value),
    timestamp = ymd_hm(paste(CollectionDateKeyValue, CollectionTimeOfDayKeyValue))
  ) %>%
  merge(baseline_platelet, by='EncounterKey', all.x=T) %>%
  merge(df_timestamp, by='EncounterKey', all.x=T) %>%
  arrange(timestamp) %>%
  filter(timestamp > begin_timestamp & timestamp < end_timestamp & NumericValue < 100 & baseline_platelet/2 >= NumericValue) %>%
  mutate(ESOFA_PLATELET = 1) %>%
  select(EncounterKey, timestamp, ESOFA_PLATELET)
```
