```{r}
nftools::nf_library(c(
  'lubridate',
  'tidyverse',
  'data.table',
  'glue'
))

base <- '~/src/Sepsis-R01/Data_All_2023-12-21/'

dx <- fread(paste0(base, 'dx_all.csv'))
ed_keys <- fread(paste0(base, 'ed_keys.csv'))
enc_dx <- fread(paste0(base, 'enc_dx.csv'))
flowsheet <- fread(paste0(base, 'flowsheet.csv'))
hosp_adm <- fread(paste0(base, 'hosp_adm.csv'))
old_labs <- fread(paste0(base, 'labs.csv'))
pt <- fread(paste0(base, 'pt.csv')) %>%
  mutate(
    has_birthdate = ifelse(is.na(BirthDate), F, T)
  )
durationdim <- fread(paste0('~/src/Sepsis-R01/', 'durationdim.csv')) %>% 
  select(DurationKey, Years) %>%
  rename(DurationAge=Years)
ed <- fread(paste0(base, 'ed.csv')) %>%
  merge(durationdim, by.x = 'AgeKey', by.y='DurationKey', all.x=T)

# Sort by timestamp
flowsheet <- flowsheet %>%
  mutate(timestamp = ymd_hm(paste(DateKeyValue, TimeOfDayKeyValue))) %>%
  dplyr::filter(EncounterKey %in% ed_keys$EncounterKey)

ed <- ed %>%
  mutate(ed_timestamp = ymd_hm(paste(ArrivalDateKeyValue, ArrivalTimeOfDayKeyValue)))

# Load supporting files
base <- '~/src/Sepsis-R01/Data_All_2024-01-31/'
#base <- '~/src/Sepsis-R01/Data_All_2_2023-12-27/'
death <- fread(paste0(base, 'death.csv'))
#ed <- fread(paste0(base, 'ed.csv'))
enc_keys <- fread(paste0(base, 'enc_keys.csv'))
flowsheet_assess <- fread(paste0(base, 'flowsheet_assess.csv'))
flowsheet_resp <- fread(paste0(base, 'flowsheet_resp.csv'))
#hosp_adm <- fread(paste0(base, 'hosp_adm.csv'))
icu_reg <- fread(paste0(base, 'icu_reg.csv'))
insurance <- fread(paste0(base, 'insurance.csv'))
labs <- fread(paste0(base, 'labs.csv'))
med_ab <- fread(paste0(base, 'med_ab.csv'))
med_vp <- fread(paste0(base, 'med_vp.csv'))

#pt <- fread(paste0(base, 'pt.csv'))
pt_keys <- fread(paste0(base, 'pt_keys.csv'))

# Sort by collection time, not result time
labs <- labs %>%
  mutate(timestamp = ymd_hm(paste(CollectionDateKeyValue, CollectionTimeOfDayKeyValue)))

# Load imaging files
base <- '~/src/Sepsis-R01/Imaging_2024-05-11/'
df_imaging_all_proc <- fread(paste0(base, 'img_all_proc.csv'))
df_imaging_lung <- fread(paste0(base, 'img_lung.csv')) %>%
  mutate(timestamp = ymd_hms(OrderingInstant))
df_imaging_lung_proc <- fread(paste0(base, 'img_lung_proc.csv'))
```

```{r WBC}
wbc <- labs %>%
  mutate(
    parsed_value = parse_number(Value),
    is_na_coercion = is.na(parsed_value)
  )
print(wbc %>% ungroup %>% filter(is_na_coercion) %>% count(Value))

wbc <- wbc %>%
  mutate(NumericValue = as.numeric(Value)) %>%
  filter(ComponentName== 'WBC Count' & !is.na(NumericValue)) %>%
  group_by(EncounterKey) %>%
  arrange(timestamp) %>%
  slice(1) %>%
  merge(ed %>% select(EncounterKey, ed_timestamp), by='EncounterKey', all.x=T) %>%
  mutate(
    TWO_HOURS = ifelse(difftime(timestamp, ed_timestamp, units = "hours") < 2, 1, 0),
    SIRS_WBC = ifelse(TWO_HOURS & (NumericValue > 12 | NumericValue < 4), 1, 0)
  ) %>%
  rename(wbc = NumericValue)

wbc %>% ungroup %>% count(SIRS_WBC)
```

```{r Temperature}
temp <- flowsheet %>%
  filter(FlowsheetRowDisplayName == 'Temp' & !is.na(NumericValue) & Accepted == 1) %>%
  mutate(temp = ifelse(NumericValue > 42, (NumericValue - 32) * 5/9, NumericValue)) %>%
  filter(temp >= 32 & temp <= 42) %>%
  group_by(EncounterKey) %>%
  arrange(timestamp) %>%
  slice(1) %>%
  merge(ed %>% select(EncounterKey, ed_timestamp), by='EncounterKey', all.x=T) %>%
  mutate(
#    TEMP_TWO_HOURS = ifelse(difftime(timestamp, ed_timestamp, units = "hours") < 2, 1, 0),
#    SIRS_TEMP = ifelse(TEMP_TWO_HOURS & (temp > 38 | temp < 36), 1, 0)
    SIRS_TEMP = ifelse(temp > 38 | temp < 36, 1, 0)
  )

temp %>% ungroup %>% count(SIRS_TEMP)
```

```{r Imaging}
df_imaging_lung <- df_imaging_lung %>%
  group_by(EncounterKey) %>%
  arrange(timestamp) %>%
  slice(1) %>%
  merge(ed %>% select(EncounterKey, ed_timestamp), by='EncounterKey', all.x=T) %>%
  mutate(
    FOUR_HOURS = ifelse(difftime(timestamp, ed_timestamp, units = "hours") < 4, 1, 0),
    IMAGING = ifelse(FOUR_HOURS == 1, 1, 0)
  )
```

```{r}
df <- ed_keys %>%
  merge(wbc %>% select(PatientDurableKey, EncounterKey, wbc, TWO_HOURS, SIRS_WBC, timestamp), by='EncounterKey', all.x=T, suffixes=c('_ED', '_WBC')) %>%
  merge(temp %>% select(PatientDurableKey, EncounterKey, temp, SIRS_TEMP, timestamp), by='EncounterKey', all.x=T, suffixes=c('', '_TEMP')) %>%
  mutate(
    analyzed = ifelse(!is.na(wbc), T, ifelse(!is.na(temp), T, F)),
    summation = rowSums(select(., contains('SIRS')), na.rm = TRUE),
    include = ifelse(summation > 0, T, F)
    )

write.csv(df, file='inclusion.csv')
```

```{r Define Sepsis ICD Codes}
icd_codes <- c(dx %>% 
                 filter(grepl('sepsis|septic', icd_l3_name, ignore.case=T)) %>% 
                 pull(icd_value) %>% 
                 unique,
               dx %>% 
                 filter(grepl('sepsis|septic', icd_l4_name, ignore.case=T)) %>% 
                 pull(icd_value) %>% 
                 unique,
               dx %>% 
                 filter(grepl('sepsis|septic', icd_l5_name, ignore.case=T)) %>%
                 pull(icd_value) %>% 
                 unique,
               dx %>% 
                 filter(grepl('sepsis|septic', icd_l6_name, ignore.case=T)) %>%
                 pull(icd_value) %>% 
                 unique
) %>% unique %>% sort

shock_icd_codes <- c(
  dx %>% 
    filter(grepl('sepsis|septic', icd_l3_name, ignore.case=T) & 
             grepl('shock', icd_l3_name, ignore.case=T) & 
             !grepl('without', icd_l3_name, ignore.case=T)) %>% 
    pull(icd_value) %>% 
    unique,
  dx %>% 
    filter(grepl('sepsis|septic', icd_l4_name, ignore.case=T) & 
             grepl('shock', icd_l4_name, ignore.case=T) & 
             !grepl('without', icd_l4_name, ignore.case=T)) %>% 
    pull(icd_value) %>% 
    unique,
  dx %>% 
    filter(grepl('sepsis|septic', icd_l5_name, ignore.case=T) & 
             grepl('shock', icd_l5_name, ignore.case=T) & 
             !grepl('without', icd_l5_name, ignore.case=T)) %>% 
    pull(icd_value) %>% 
    unique,
  dx %>% 
    filter(grepl('sepsis|septic', icd_l6_name, ignore.case=T) & 
             grepl('shock', icd_l6_name, ignore.case=T) & 
             !grepl('without', icd_l6_name, ignore.case=T)) %>% 
    pull(icd_value) %>% 
    unique,
  dx %>% 
    filter(grepl('sepsis|septic', icd_l7_name, ignore.case=T) & 
             grepl('shock', icd_l7_name, ignore.case=T) & 
             !grepl('without', icd_l7_name, ignore.case=T)) %>% 
    pull(icd_value) %>% 
    unique
) %>% unique %>% sort
```


```{r}
dx_trim <- dx %>%
  select(
    'EncounterKey',
    'icd_value'
  ) %>%
  mutate(
    sepsis = ifelse(icd_value %in% icd_codes, 1, 0),
    septic_shock = ifelse(icd_value %in% shock_icd_codes, 1, 0)
  ) %>%
  filter(EncounterKey %in% ed_keys$EncounterKey) %>%
  group_by(EncounterKey) %>%
  summarize(
    sepsis = max(sepsis),
    septic_shock = max(septic_shock)
  ) %>%
  mutate(
    sepsis = ifelse(sepsis == 1, T, F),
    septic_shock = ifelse(septic_shock == 1, T, F),
  ) %>%
  ungroup

pt_trim <- pt %>%
  select(
    'Sex',
    'PreferredLanguage',
    'FirstRace',
    'SecondRace',
    'BirthDate',
    'has_birthdate',
    'PatientDurableKey',
    'Status',
    'MaritalStatus',
    'Religion',
    'SmokingStatus',
    'PrimaryFinancialClass',
    'HighestLevelOfEducation',
    'SexAssignedAtBirth',
    'GenderIdentity',
    'SexualOrientation',
    'UCSFDerivedRaceEthnicity_X'
  ) %>%
  distinct(.keep_all = TRUE) %>%
  group_by(PatientDurableKey) %>%
  slice(1) %>%
  ungroup

# replace any value with the string 'Unknown' or 'Decline' ignoring case with NA 
pt_trim <- pt_trim %>%
  mutate_if(is.character, 
    list(~ ifelse(grepl("Unknown|Decline", ., ignore.case = T), NA, .))
  )

tabledf <- df %>%
  merge(dx_trim, by='EncounterKey', all.x = T) %>%
  merge(ed %>% dplyr::select(EncounterKey, DurationAge), by='EncounterKey', all.x = T) %>% 
  merge(pt_trim, by='PatientDurableKey', all.x = T) %>%
  mutate(
    processed_birthdate = lubridate::ymd(BirthDate),
    age_days = as.numeric(difftime(lubridate::ymd(ArrivalDateKeyValue), processed_birthdate, units='days')),
    age = age_days/365.25,
    age_delta = age - DurationAge,
    SIRS_WBC = as.factor(SIRS_WBC),
    SIRS_TEMP = as.factor(SIRS_TEMP),
  ) %>%
  mutate_if(is.character, 
    list(~ ifelse(grepl("^\\*|^$|^Missing$", .), NA, .))
  ) %>%
  merge(df_imaging_lung %>% dplyr::select(EncounterKey, IMAGING, FirstProcedureName), by='EncounterKey', all.x = T) %>%
  mutate(
    IMAGING = ifelse(is.na(IMAGING), 0, IMAGING),
    include_imaging = ifelse(include == 1 & IMAGING == 1, 
                             'Suspected Sepsis with Imaging',
                             ifelse(include == 1,
                                    'Suspected Sepsis No Imaging',
                                    'No Suspected Sepsis'
                                    )
    )
  )

columns <- c(
#    'Status',
#    'age',
    'Sex',
#    'PreferredLanguage',
    'FirstRace',
#    'SecondRace',
#    'analyzed',
#    'SIRS_WBC',
#    'SIRS_TEMP',
    'DurationAge',
#    'age_delta',
#    'has_birthdate',
#    'SexAssignedAtBirth',
#    'GenderIdentity',
#    'MaritalStatus',
#    'Religion',
#    'SmokingStatus',
#    'PrimaryFinancialClass',
#    'SexualOrientation',
    'wbc',
    'temp',
    'sepsis',
    'septic_shock'
)

nftools::nf_table1(columns, 'include', tabledf, calculate_pvalue = F)
nftools::nf_table1(columns, 'include_imaging', tabledf, calculate_pvalue = F)

tabledf$year <- tabledf$ArrivalDateKeyValue%>% year
nftools::nf_table1(columns, 'year', tabledf, calculate_pvalue = F)
```
```{r Summary stats for R01 Writeup}
# Number of visits per year
tabledf %>% count(year) %>% pull(n) %>% summary
```
```{r Determine Flag Timestamp}
df_timestamp <- df %>%
  filter(include == T) %>%
  rename(timestamp_WBC=timestamp) %>%
  mutate(timestamp=ifelse(is.na(SIRS_TEMP) | SIRS_TEMP == 0,
                          timestamp_WBC,
                          ifelse(is.na(SIRS_WBC) | SIRS_WBC == 0,
                                 timestamp_TEMP,
                                 ifelse(timestamp_WBC < timestamp_TEMP,
                                        timestamp_WBC,
                                        timestamp_TEMP)))) %>%
  mutate(timestamp = as_datetime(ymd_hms("1970-01-01 00:00:00") + seconds(timestamp))) %>%
  select(contains('Key'), contains('timestamp'), contains('SIRS')) %>%
  mutate(
    begin_timestamp = timestamp + hours(48) - hours(4),
    end_timestamp = timestamp + hours(48) + hours(4),
    midpoint_timestamp = timestamp + hours(48)
  ) %>%
  rename(
    inclusion_timestamp=timestamp
  )

df_timestamp <- df_timestamp %>% select(
  'EncounterKey',
  'inclusion_timestamp',
  'begin_timestamp', 
  'end_timestamp',
  'midpoint_timestamp'
)

df <- df %>%
  merge(df_timestamp, by='EncounterKey', all.x=T)
```

```{r eSOFA Vasopressor}
# Any vasopressor administration between ED encounter and 48 hours
esofa_pressors <- med_vp %>%
  filter(EncounterKey %in% ed_keys$EncounterKey) %>%
  mutate(timestamp = ymd_hm(paste(AdministrationDateKeyValue, AdministrationTimeOfDayKeyValue))) %>%
  merge(df_timestamp, by='EncounterKey', all.x=T) %>%
#  filter(timestamp > begin_timestamp & timestamp < end_timestamp) %>%
  filter(timestamp < midpoint_timestamp) %>%
  group_by(EncounterKey) %>%
  slice(1) %>%
  mutate(ESOFA_PRESSOR = 1) %>%
  select(EncounterKey, timestamp, ESOFA_PRESSOR)
```

> icu_reg %>% count(IntubationDays)
> resp %>% count(FlowsheetRowName) %>% as.data.table %>% arrange(-n)
                                FlowsheetRowName      n
 1:                                       R FIO2 692846
 2:                         R RESP PEEP/CPAP SET  96036 #Good
 3:                               R RT VENT MODE  92240 #Good
 4:                        R RT MODE/BREATH TYPE  91142 #Good but with DEID columns
 5:      R RT VENTILATOR RESPIRATORY RATE  TOTAL  90097
 6:                   RT MAP MEASURED_IP_CD_UCSF  89282
 7:                       R RT (ADULT) VENT TYPE  84673
 8:    R RT VENTILATOR MINUTE VOLUME MEASURED V2  70198
 9: R RT (ADULT) VENTILATOR TIDAL VOLUME EXHALED  58667
10:   R RT (ADULT) VENTILATOR I:E RATIO MEASURED  56879
11:                                R RESP RR SET  53608
12:              R OXYGEN SATURATION INDEX (OSI)  51269
13:                                     R FIO2 2   9678

```{r eSOFA Mechanical Ventilation}
# Any vasopressor administration between ED encounter and 48 hours
esofa_mv <- flowsheet_resp %>%
  filter(EncounterKey %in% ed_keys$EncounterKey) %>%
  filter(FlowsheetRowName %in% c('R RESP PEEP/CPAP SET', 'R RT VENT MODE', 'R RT MODE/BREATH TYPE')) %>%
  mutate(timestamp = ymd_hm(paste(DateKeyValue, TimeOfDayKeyValue))) %>%
  merge(df_timestamp, by='EncounterKey', all.x=T) %>%
#  filter(timestamp > begin_timestamp & timestamp < end_timestamp) %>%
  filter(timestamp < midpoint_timestamp) %>%
  arrange(timestamp) %>%
  group_by(EncounterKey) %>%
  slice(1) %>%
  mutate(ESOFA_MV = 1) %>%
  select(EncounterKey, timestamp, ESOFA_MV)
```

> labs %>% count(ComponentName)
                             ComponentName      n
 1:                      Bilirubin, Direct  12337
 2:                       Bilirubin, Total 139478
 3:                             Creatinine 352940
 4:                 Creatinine Whole Blood     35
 5:                      Creatinine, Serum     60
 6:                             Hemoglobin 334015
 7: Hemoglobin, Plasma, Alternative Method    182
 8:                        Lactate, plasma  15653
 9:                   Lactate, whole blood 117056
10:                                    PO2 161882
11:                            PO2, Venous     42
12:                 POCT, Total Hemoglobin    598
13:                         Platelet Count 336999
14:                       Total Hemoglobin      5
15:                              WBC Count 333037

```{r eSOFA Lactate}
esofa_lactate <- labs %>%
  filter(EncounterKey %in% ed_keys$EncounterKey) %>%
  filter(ComponentName %in% c('Lactate, plasma')) %>%
  mutate(timestamp = ymd_hm(paste(CollectionDateKeyValue, CollectionTimeOfDayKeyValue))) %>%
  merge(df_timestamp, by='EncounterKey', all.x=T) %>%
  filter(timestamp > begin_timestamp & timestamp < end_timestamp) %>%
  arrange(timestamp) %>%
  filter(Value >= 2) %>%
  mutate(ESOFA_LACTATE = 1) %>%
  select(EncounterKey, timestamp, ESOFA_LACTATE)
```

```{r Doubling of Creatinine}
baseline_creatinine <- labs %>%
  filter(EncounterKey %in% ed_keys$EncounterKey & ComponentName %in% c('Creatinine', 'Creatinine, Serum')) %>%
  mutate(
    NumericValue = as.numeric(Value),
    timestamp = ymd_hm(paste(CollectionDateKeyValue, CollectionTimeOfDayKeyValue))
  ) %>%
  arrange(timestamp) %>%
  group_by(EncounterKey) %>%
  slice(1) %>%
  select(EncounterKey, NumericValue) %>%
  rename(baseline_creatinine=NumericValue) %>%
  ungroup

esofa_creatinine <- labs %>%
  filter(EncounterKey %in% ed_keys$EncounterKey & ComponentName %in% c('Creatinine', 'Creatinine, Serum')) %>%
  mutate(
    NumericValue = as.numeric(Value),
    timestamp = ymd_hm(paste(CollectionDateKeyValue, CollectionTimeOfDayKeyValue))
  ) %>%
  merge(baseline_creatinine, by='EncounterKey', all.x=T) %>%
  merge(df_timestamp, by='EncounterKey', all.x=T) %>%
  arrange(timestamp) %>%
  filter(timestamp > begin_timestamp & timestamp < end_timestamp & NumericValue >= baseline_creatinine * 2) %>%
  mutate(ESOFA_CREATININE= 1) %>%
  select(EncounterKey, timestamp, ESOFA_CREATININE)
```

```{r Doubling of Bilirubin to ≥2 mg/dL}
baseline_bilirubin <- labs %>%
  filter(EncounterKey %in% ed_keys$EncounterKey & ComponentName %in% c('Bilirubin, Total')) %>%
  mutate(
    NumericValue = as.numeric(Value),
    timestamp = ymd_hm(paste(CollectionDateKeyValue, CollectionTimeOfDayKeyValue))
  ) %>%
  arrange(timestamp) %>%
  group_by(EncounterKey) %>%
  slice(1) %>%
  select(EncounterKey, NumericValue) %>%
  rename(baseline_bilirubin=NumericValue) %>%
  ungroup

esofa_bilirubin <- labs %>%
  filter(EncounterKey %in% ed_keys$EncounterKey & ComponentName %in% c('Bilirubin, Total')) %>%
  mutate(
    NumericValue = as.numeric(Value),
    timestamp = ymd_hm(paste(CollectionDateKeyValue, CollectionTimeOfDayKeyValue))
  ) %>%
  merge(baseline_bilirubin, by='EncounterKey', all.x=T) %>%
  merge(df_timestamp, by='EncounterKey', all.x=T) %>%
  arrange(timestamp) %>%
  filter(timestamp > begin_timestamp & timestamp < end_timestamp & NumericValue >= baseline_bilirubin * 2 & NumericValue >= 2) %>%
  mutate(ESOFA_BILIRUBIN = 1) %>%
  select(EncounterKey, timestamp, ESOFA_BILIRUBIN)
```

```{r ≥50% Decrease in Platelet Count to less than 100 cells/microL}
baseline_platelet <- labs %>%
  filter(EncounterKey %in% ed_keys$EncounterKey & ComponentName %in% c('Platelet Count')) %>%
  mutate(
    NumericValue = as.numeric(Value),
    timestamp = ymd_hm(paste(CollectionDateKeyValue, CollectionTimeOfDayKeyValue))
  ) %>%
  arrange(timestamp) %>%
  group_by(EncounterKey) %>%
  slice(1) %>%
  select(EncounterKey, NumericValue) %>%
  filter(NumericValue >= 100) %>% # Filter out baseline values < 100
  rename(baseline_platelet=NumericValue) %>%
  ungroup

esofa_platelet <- labs %>%
  filter(EncounterKey %in% ed_keys$EncounterKey & ComponentName %in% c('Platelet Count')) %>%
  mutate(
    NumericValue = as.numeric(Value),
    timestamp = ymd_hm(paste(CollectionDateKeyValue, CollectionTimeOfDayKeyValue))
  ) %>%
  merge(baseline_platelet, by='EncounterKey', all.x=T) %>%
  merge(df_timestamp, by='EncounterKey', all.x=T) %>%
  arrange(timestamp) %>%
  filter(timestamp > begin_timestamp & timestamp < end_timestamp & NumericValue < 100 & baseline_platelet/2 >= NumericValue) %>%
  mutate(ESOFA_PLATELET = 1) %>%
  select(EncounterKey, timestamp, ESOFA_PLATELET)
```


```{r Merge eSOFA}
# merge each esofa_ dataframe, selecting only EncounterKey and the ESOFA_ column to be merged in
esofa <- df_timestamp %>%
  merge(esofa_pressors %>% select(EncounterKey, ESOFA_PRESSOR), by='EncounterKey', all.x=T) %>%
  merge(esofa_mv %>% select(EncounterKey, ESOFA_MV), by='EncounterKey', all.x=T) %>%
  merge(esofa_lactate %>% select(EncounterKey, ESOFA_LACTATE), by='EncounterKey', all.x=T) %>%
  merge(esofa_creatinine %>% select(EncounterKey, ESOFA_CREATININE), by='EncounterKey', all.x=T) %>%
  merge(esofa_bilirubin %>% select(EncounterKey, ESOFA_BILIRUBIN), by='EncounterKey', all.x=T) %>%
  merge(esofa_platelet %>% select(EncounterKey, ESOFA_PLATELET), by='EncounterKey', all.x=T) %>%
  mutate(
    ESOFA_COUNT = rowSums(select(., starts_with('ESOFA')), na.rm=T),
    ESOFA_BINARY  = ifelse(ESOFA_COUNT > 0, 1, 0)
  )
```

```{r Create Epochs}
df_timestamp <- df_timestamp %>%
  mutate(
    epoch_1_begin = inclusion_timestamp,
    epoch_1_end = inclusion_timestamp + minutes(30),
    epoch_2_begin = inclusion_timestamp + minutes(30),
    epoch_2_end = inclusion_timestamp + minutes(60),
    epoch_3_begin = inclusion_timestamp + minutes(60),
    epoch_3_end = inclusion_timestamp + minutes(90),
    epoch_4_begin = inclusion_timestamp + minutes(90),
    epoch_4_end = inclusion_timestamp + minutes(120),
    epoch_5_begin = inclusion_timestamp + minutes(120),
    epoch_5_end = inclusion_timestamp + minutes(150),
    epoch_6_begin = inclusion_timestamp + minutes(150),
    epoch_6_end = inclusion_timestamp + minutes(180),
    epoch_7_begin = inclusion_timestamp + minutes(180),
    epoch_7_end = inclusion_timestamp + minutes(210),
    epoch_8_begin = inclusion_timestamp + minutes(210),
    epoch_8_end = inclusion_timestamp + minutes(240),
    epoch_9_begin = inclusion_timestamp + minutes(240),
    epoch_9_end = inclusion_timestamp + minutes(270),
    epoch_10_begin = inclusion_timestamp + minutes(270),
    epoch_10_end = inclusion_timestamp + minutes(300),
    epoch_11_begin = inclusion_timestamp + minutes(300),
    epoch_11_end = inclusion_timestamp + minutes(330),
    epoch_12_begin = inclusion_timestamp + minutes(330),
    epoch_12_end = inclusion_timestamp + minutes(360),
    epoch_13_begin = inclusion_timestamp + minutes(360),
    epoch_13_end = inclusion_timestamp + minutes(390),
    epoch_14_begin = inclusion_timestamp + minutes(390),
    epoch_14_end = inclusion_timestamp + minutes(420),
    epoch_15_begin = inclusion_timestamp + minutes(420),
    epoch_15_end = inclusion_timestamp + minutes(450),
    epoch_16_begin = inclusion_timestamp + minutes(450),
    epoch_16_end = inclusion_timestamp + minutes(480),
    epoch_17_begin = inclusion_timestamp + minutes(480),
    epoch_17_end = inclusion_timestamp + minutes(510),
    epoch_18_begin = inclusion_timestamp + minutes(510),
    epoch_18_end = inclusion_timestamp + minutes(540),
    epoch_19_begin = inclusion_timestamp + minutes(540),
    epoch_19_end = inclusion_timestamp + minutes(570),
    epoch_20_begin = inclusion_timestamp + minutes(570),
    epoch_20_end = inclusion_timestamp + minutes(600),
    epoch_21_begin = inclusion_timestamp + minutes(600),
    epoch_21_end = inclusion_timestamp + minutes(630),
    epoch_22_begin = inclusion_timestamp + minutes(630),
    epoch_22_end = inclusion_timestamp + minutes(660),
    epoch_23_begin = inclusion_timestamp + minutes(660),
    epoch_23_end = inclusion_timestamp + minutes(690),
    epoch_24_begin = inclusion_timestamp + minutes(690),
    epoch_24_end = inclusion_timestamp + minutes(720)
  ) %>%
  merge(ed %>% dplyr::select(EncounterKey, DurationAge), by='EncounterKey', all.x = T)
  
generate_lab_epochs <- function(df, component_name, label) {
  output <- df %>%
    filter(ComponentName %in% component_name) %>%
    filter(EncounterKey %in% df_timestamp$EncounterKey) %>%
    mutate(
      NumericValue = as.numeric(Value),
      timestamp = ymd_hm(paste(CollectionDateKeyValue, CollectionTimeOfDayKeyValue))
    ) %>%
    merge(df_timestamp, by='EncounterKey', all.x=T) %>%
    mutate(
      epoch_label = ifelse(timestamp >= epoch_1_begin & timestamp < epoch_1_end,
                                1,
                                ifelse(timestamp >= epoch_2_begin & timestamp < epoch_2_end,
                                       2,
                                       ifelse(timestamp >= epoch_3_begin & timestamp < epoch_3_end,
                                              3,
                                              ifelse(timestamp >= epoch_4_begin & timestamp < epoch_4_end,
                                                     4,
                                                     ifelse(timestamp >= epoch_5_begin & timestamp < epoch_5_end,
                                                            5,
                                                            ifelse(timestamp >= epoch_6_begin & timestamp < epoch_6_end,
                                                                   6,
                                                                   ifelse(timestamp >= epoch_7_begin & timestamp < epoch_7_end,
                                                                          7,
                                                                          ifelse(timestamp >= epoch_8_begin & timestamp < epoch_8_end,
                                                                                 8,
                                                                                 ifelse(timestamp >= epoch_9_begin & timestamp < epoch_9_end,
                                                                                        9,
                                                                                        ifelse(timestamp >= epoch_10_begin & timestamp < epoch_10_end,
                                                                                               10,
                                                                                               ifelse(timestamp >= epoch_11_begin & timestamp < epoch_11_end,
                                                                                                      11,
                                                                                                      ifelse(timestamp >= epoch_12_begin & timestamp < epoch_12_end,
                                                                                                             12,
                                                                                                             ifelse(timestamp >= epoch_13_begin & timestamp < epoch_13_end,
                                                                                                                    13,
                                                                                                                    ifelse(timestamp >= epoch_14_begin & timestamp < epoch_14_end,
                                                                                                                           14,
                                                                                                                           ifelse(timestamp >= epoch_15_begin & timestamp < epoch_15_end,
                                                                                                                                  15,
                                                                                                                                  ifelse(timestamp >= epoch_16_begin & timestamp < epoch_16_end,
                                                                                                                                         16,
                                                                                                                                         ifelse(timestamp >= epoch_17_begin & timestamp < epoch_17_end,
                                                                                                                                                17,
                                                                                                                                                ifelse(timestamp >= epoch_18_begin & timestamp < epoch_18_end,
                                                                                                                                                       18,
                                                                                                                                                       ifelse(timestamp >= epoch_19_begin & timestamp < epoch_19_end,
                                                                                                                                                              19,
                                                                                                                                                              ifelse(timestamp >= epoch_20_begin & timestamp < epoch_20_end,
                                                                                                                                                                     20,
                                                                                                                                                                     ifelse(timestamp >= epoch_21_begin & timestamp < epoch_21_end,
                                                                                                                                                                            21,
                                                                                                                                                                            ifelse(timestamp >= epoch_22_begin & timestamp < epoch_22_end,
                                                                                                                                                                                   22,
                                                                                                                                                                                   ifelse(timestamp >= epoch_23_begin & timestamp < epoch_23_end,
                                                                                                                                                                                          23,
                                                                                                                                                                                          ifelse(timestamp >= epoch_24_begin & timestamp < epoch_24_end,
                                                                                                                                                                                                 24,
                                                                                                                                                                                                 NA
                                                                                                                                                                                                 ))))))))))))))))))))))))
    ) %>%
    dcast(EncounterKey ~ epoch_label, value.var = "NumericValue", fun.aggregate = max) %>%
    select(-'NA')
  
  colnames(output) <- c("EncounterKey", paste0("epoch_", colnames(output)[-1], "_", label))
  output[output == -Inf] <- NA
  return(output)
}

epochs_creatinine <- generate_lab_epochs(labs, c('Creatinine', 'Creatinine, Serum'), 'creatinine')
epochs_lactate <- generate_lab_epochs(labs, c('Lactate, plasma'), 'lactate')
epochs_hemoglobin <- generate_lab_epochs(labs, c('Hemoglobin'), 'hgb')
epochs_bilirubin <- generate_lab_epochs(labs, c('Bilirubin, Total'), 'bili')
epochs_platelet <- generate_lab_epochs(labs, c('Platelet Count'), 'plt')
epochs_wbc <- generate_lab_epochs(labs, c('WBC Count'), 'wbc')
epochs_po2 <- generate_lab_epochs(labs, c('PO2'), 'po2')

generate_vitals_epochs <- function(df, component_name, label) {
  output <- df %>%
    filter(FlowsheetRowDisplayName %in% component_name) %>%
    filter(EncounterKey %in% df_timestamp$EncounterKey) %>%
    mutate(
      systolic = as.numeric(str_extract(Value, "^[0-9]+")),
      # set diastolic to the second half of the string in Value after the /
      diastolic = as.numeric(str_extract(Value, "(?<=/)[0-9]+")),
      # calculate the map from systolic and diastolic
      map = (systolic + 2*diastolic) / 3,
      NumericValue = ifelse(FlowsheetRowDisplayName == 'BP', map, as.numeric(Value)),
      timestamp = ymd_hm(paste(DateKeyValue, TimeOfDayKeyValue)),
    ) %>%
    merge(df_timestamp, by='EncounterKey', all.x=T) %>%
    mutate(
      epoch_label = ifelse(timestamp >= epoch_1_begin & timestamp < epoch_1_end,
                                1,
                                ifelse(timestamp >= epoch_2_begin & timestamp < epoch_2_end,
                                       2,
                                       ifelse(timestamp >= epoch_3_begin & timestamp < epoch_3_end,
                                              3,
                                              ifelse(timestamp >= epoch_4_begin & timestamp < epoch_4_end,
                                                     4,
                                                     ifelse(timestamp >= epoch_5_begin & timestamp < epoch_5_end,
                                                            5,
                                                            ifelse(timestamp >= epoch_6_begin & timestamp < epoch_6_end,
                                                                   6,
                                                                   ifelse(timestamp >= epoch_7_begin & timestamp < epoch_7_end,
                                                                          7,
                                                                          ifelse(timestamp >= epoch_8_begin & timestamp < epoch_8_end,
                                                                                 8,
                                                                                 ifelse(timestamp >= epoch_9_begin & timestamp < epoch_9_end,
                                                                                        9,
                                                                                        ifelse(timestamp >= epoch_10_begin & timestamp < epoch_10_end,
                                                                                               10,
                                                                                               ifelse(timestamp >= epoch_11_begin & timestamp < epoch_11_end,
                                                                                                      11,
                                                                                                      ifelse(timestamp >= epoch_12_begin & timestamp < epoch_12_end,
                                                                                                             12,
                                                                                                             ifelse(timestamp >= epoch_13_begin & timestamp < epoch_13_end,
                                                                                                                    13,
                                                                                                                    ifelse(timestamp >= epoch_14_begin & timestamp < epoch_14_end,
                                                                                                                           14,
                                                                                                                           ifelse(timestamp >= epoch_15_begin & timestamp < epoch_15_end,
                                                                                                                                  15,
                                                                                                                                  ifelse(timestamp >= epoch_16_begin & timestamp < epoch_16_end,
                                                                                                                                         16,
                                                                                                                                         ifelse(timestamp >= epoch_17_begin & timestamp < epoch_17_end,
                                                                                                                                                17,
                                                                                                                                                ifelse(timestamp >= epoch_18_begin & timestamp < epoch_18_end,
                                                                                                                                                       18,
                                                                                                                                                       ifelse(timestamp >= epoch_19_begin & timestamp < epoch_19_end,
                                                                                                                                                              19,
                                                                                                                                                              ifelse(timestamp >= epoch_20_begin & timestamp < epoch_20_end,
                                                                                                                                                                     20,
                                                                                                                                                                     ifelse(timestamp >= epoch_21_begin & timestamp < epoch_21_end,
                                                                                                                                                                            21,
                                                                                                                                                                            ifelse(timestamp >= epoch_22_begin & timestamp < epoch_22_end,
                                                                                                                                                                                   22,
                                                                                                                                                                                   ifelse(timestamp >= epoch_23_begin & timestamp < epoch_23_end,
                                                                                                                                                                                          23,
                                                                                                                                                                                          ifelse(timestamp >= epoch_24_begin & timestamp < epoch_24_end,
                                                                                                                                                                                                 24,
                                                                                                                                                                                                 NA
                                                                                                                                                                                                 ))))))))))))))))))))))))
    ) %>%
    dcast(EncounterKey ~ epoch_label, value.var = "NumericValue", fun.aggregate = max) %>%
    select(-'NA')
  
  colnames(output) <- c("EncounterKey", paste0("epoch_", colnames(output)[-1], "_", label))
  output[output == -Inf] <- NA
  return(output)
}

epochs_pulse <- generate_vitals_epochs(flowsheet_assess, c('Pulse', 'Pulse - Palpated/Pleth'), 'pulse')
epochs_resp <- generate_vitals_epochs(flowsheet_assess, c('Resp'), 'resp')
epochs_temp <- generate_vitals_epochs(flowsheet_assess, c('Temp'), 'temp')
epochs_map <- generate_vitals_epochs(flowsheet_assess, c('MAP (mmHg)', 'Arterial Line MAP (mmHg)', 'BP'), 'map')
epochs_gcse <- generate_vitals_epochs(flowsheet_assess, c('Best Eye Response'), 'gcs_eye')
epochs_gcsv <- generate_vitals_epochs(flowsheet_assess, c('Best Verbal Response'), 'gcs_verbal')
epochs_gcsm <- generate_vitals_epochs(flowsheet_assess, c('Best Motor Response (upper extremities)'), 'gcs_motor')
epochs_weight <- generate_vitals_epochs(flowsheet_assess, c('Weight'), 'weight')

generate_mar_epochs <- function(df, component_name, label) {
  output <- df %>%
#    filter(FlowsheetRowDisplayName == component_name) %>%
    filter(EncounterKey %in% df_timestamp$EncounterKey) %>%
    filter(AdministrationAction %in% c('Given', 
                                       'Given During Downtime', 
                                       'Given by Other', 
                                       #'MAR Unhold', 
                                       'New Bag', 
                                       'New Syringe', 
                                       'Started During Downtime')) %>%
    mutate(
      #NumericValue = as.numeric(Value),
      timestamp = ymd_hm(paste(AdministrationDateKeyValue, AdministrationTimeOfDayKeyValue)),
    ) %>%
    arrange(EncounterKey, timestamp) %>%
    group_by(EncounterKey) %>%
    slice(1) %>%
    ungroup() %>%
    merge(df_timestamp, by='EncounterKey', all.x=T) %>%
    mutate(
      epoch_label = ifelse(timestamp >= epoch_1_begin & timestamp < epoch_1_end,
                                1,
                                ifelse(timestamp >= epoch_2_begin & timestamp < epoch_2_end,
                                       2,
                                       ifelse(timestamp >= epoch_3_begin & timestamp < epoch_3_end,
                                              3,
                                              ifelse(timestamp >= epoch_4_begin & timestamp < epoch_4_end,
                                                     4,
                                                     ifelse(timestamp >= epoch_5_begin & timestamp < epoch_5_end,
                                                            5,
                                                            ifelse(timestamp >= epoch_6_begin & timestamp < epoch_6_end,
                                                                   6,
                                                                   ifelse(timestamp >= epoch_7_begin & timestamp < epoch_7_end,
                                                                          7,
                                                                          ifelse(timestamp >= epoch_8_begin & timestamp < epoch_8_end,
                                                                                 8,
                                                                                 ifelse(timestamp >= epoch_9_begin & timestamp < epoch_9_end,
                                                                                        9,
                                                                                        ifelse(timestamp >= epoch_10_begin & timestamp < epoch_10_end,
                                                                                               10,
                                                                                               ifelse(timestamp >= epoch_11_begin & timestamp < epoch_11_end,
                                                                                                      11,
                                                                                                      ifelse(timestamp >= epoch_12_begin & timestamp < epoch_12_end,
                                                                                                             12,
                                                                                                             ifelse(timestamp >= epoch_13_begin & timestamp < epoch_13_end,
                                                                                                                    13,
                                                                                                                    ifelse(timestamp >= epoch_14_begin & timestamp < epoch_14_end,
                                                                                                                           14,
                                                                                                                           ifelse(timestamp >= epoch_15_begin & timestamp < epoch_15_end,
                                                                                                                                  15,
                                                                                                                                  ifelse(timestamp >= epoch_16_begin & timestamp < epoch_16_end,
                                                                                                                                         16,
                                                                                                                                         ifelse(timestamp >= epoch_17_begin & timestamp < epoch_17_end,
                                                                                                                                                17,
                                                                                                                                                ifelse(timestamp >= epoch_18_begin & timestamp < epoch_18_end,
                                                                                                                                                       18,
                                                                                                                                                       ifelse(timestamp >= epoch_19_begin & timestamp < epoch_19_end,
                                                                                                                                                              19,
                                                                                                                                                              ifelse(timestamp >= epoch_20_begin & timestamp < epoch_20_end,
                                                                                                                                                                     20,
                                                                                                                                                                     ifelse(timestamp >= epoch_21_begin & timestamp < epoch_21_end,
                                                                                                                                                                            21,
                                                                                                                                                                            ifelse(timestamp >= epoch_22_begin & timestamp < epoch_22_end,
                                                                                                                                                                                   22,
                                                                                                                                                                                   ifelse(timestamp >= epoch_23_begin & timestamp < epoch_23_end,
                                                                                                                                                                                          23,
                                                                                                                                                                                          ifelse(timestamp >= epoch_24_begin & timestamp < epoch_24_end,
                                                                                                                                                                                                 24,
                                                                                                                                                                                                 NA
                                                                                                                                                                                                 ))))))))))))))))))))))))
    ) %>%
    as.data.table %>%
    dcast(EncounterKey ~ epoch_label, value.var = "AdministrationAction", fun.aggregate = function(x) paste(x, collapse='_')) %>%
    select(-'NA')
  
  colnames(output) <- c("EncounterKey", paste0("epoch_", colnames(output)[-1], "_", label))
  output[output == -Inf] <- NA
  output[output == ''] <- NA
  
  return(output)
}



# restrict med_ab to the first 105 columns
epochs_abx <- generate_mar_epochs(med_ab[,1:105], 'Antibiotics', 'abx')

# get the last lab before inclusion 
get_last_lab <- function(df, component_name, label) {
  output <- df %>%
    filter(ComponentName %in% component_name) %>%
    filter(EncounterKey %in% df_timestamp$EncounterKey) %>%
    mutate(
      NumericValue = as.numeric(Value),
      timestamp = ymd_hm(paste(CollectionDateKeyValue, CollectionTimeOfDayKeyValue))
    ) %>%
    merge(df_timestamp, by='EncounterKey', all.x=T) %>%
    filter(timestamp < inclusion_timestamp) %>%
    arrange(EncounterKey, timestamp) %>%
    group_by(EncounterKey) %>%
    slice(n())
  
  output <- output %>% select(EncounterKey, NumericValue, timestamp) #, inclusion_timestamp)
  
  # rename the NumericValue column to the value of label
  colnames(output)[colnames(output) == 'NumericValue'] <- paste0('pre_inclusion_', label)
  colnames(output)[colnames(output) == 'timestamp'] <- paste0('pre_inclusion_', label, '_timestamp')
  return(output)
}

last_creatinine <- get_last_lab(labs, c('Creatinine', 'Creatinine, Serum'), 'creatinine')
last_lactate <- get_last_lab(labs, c('Lactate, plasma'), 'lactate')
last_hemoglobin <- get_last_lab(labs, c('Hemoglobin'), 'hgb')
last_bilirubin <- get_last_lab(labs, c('Bilirubin, Total'), 'bili')
last_platelet <- get_last_lab(labs, c('Platelet Count'), 'plt')
last_wbc <- get_last_lab(labs, c('WBC Count'), 'wbc')
last_po2 <- get_last_lab(labs, c('PO2'), 'po2')

# get the last vitals before inclusion 
get_last_vitals <- function(df, component_name, label) {
  output <- df %>%
    filter(FlowsheetRowDisplayName == component_name) %>%
    filter(EncounterKey %in% df_timestamp$EncounterKey) %>%
    mutate(
      systolic = as.numeric(str_extract(Value, "^[0-9]+")),
      # set diastolic to the second half of the string in Value after the /
      diastolic = as.numeric(str_extract(Value, "(?<=/)[0-9]+")),
      # calculate the map from systolic and diastolic
      map = (systolic + 2*diastolic) / 3,
      NumericValue = ifelse(FlowsheetRowDisplayName == 'BP', map, as.numeric(Value)),
      timestamp = ymd_hm(paste(DateKeyValue, TimeOfDayKeyValue)),
    ) %>%
    merge(df_timestamp, by='EncounterKey', all.x=T) %>%
    filter(timestamp < inclusion_timestamp) %>%
    arrange(EncounterKey, timestamp) %>%
    group_by(EncounterKey) %>%
    slice(n())

  output <- output %>% select(EncounterKey, NumericValue, timestamp) #, inclusion_timestamp)
  
  # rename the NumericValue column to the value of label
  colnames(output)[colnames(output) == 'NumericValue'] <- paste0('pre_inclusion_', label)
  colnames(output)[colnames(output) == 'timestamp'] <- paste0('pre_inclusion_', label, '_timestamp')
  return(output)
}

last_pulse <- get_last_vitals(flowsheet_assess, c('Pulse', 'Pulse - Palpated/Pleth'), 'pulse')
last_resp <- get_last_vitals(flowsheet_assess, c('Resp'), 'resp')
last_temp <- get_last_vitals(flowsheet_assess, c('Temp'), 'temp')
last_map <- get_last_vitals(flowsheet_assess, c('MAP (mmHg)', 'Arterial Line MAP (mmHg)', 'BP'), 'map')
last_gcse <- get_last_vitals(flowsheet_assess, c('Best Eye Response'), 'gcs_eye')
last_gcsv <- get_last_vitals(flowsheet_assess, c('Best Verbal Response'), 'gcs_verbal')
last_gcsm <- get_last_vitals(flowsheet_assess, c('Best Motor Response (upper extremities)'), 'gcs_motor')
last_weight <- get_last_vitals(flowsheet_assess, c('Weight'), 'weight')

# Merge epoch dataframes
output_df <- df_timestamp %>%
  merge(epochs_creatinine, by='EncounterKey', all.x=T) %>%
  merge(last_creatinine, by='EncounterKey', all.x=T) %>%
  merge(epochs_lactate, by='EncounterKey', all.x=T) %>%
  merge(last_lactate, by='EncounterKey', all.x=T) %>%
  merge(epochs_hemoglobin, by='EncounterKey', all.x=T) %>%
  merge(last_hemoglobin, by='EncounterKey', all.x=T) %>%
  merge(epochs_bilirubin, by='EncounterKey', all.x=T) %>%
  merge(last_bilirubin, by='EncounterKey', all.x=T) %>%
  merge(epochs_platelet, by='EncounterKey', all.x=T) %>%
  merge(last_platelet, by='EncounterKey', all.x=T) %>%
  merge(epochs_wbc, by='EncounterKey', all.x=T) %>%
  merge(last_wbc, by='EncounterKey', all.x=T) %>%
  merge(epochs_po2, by='EncounterKey', all.x=T) %>%
  merge(last_po2, by='EncounterKey', all.x=T) %>%
  merge(epochs_pulse, by='EncounterKey', all.x=T) %>%
  merge(last_pulse, by='EncounterKey', all.x=T) %>%
  merge(epochs_resp, by='EncounterKey', all.x=T) %>%
  merge(last_resp, by='EncounterKey', all.x=T) %>%
  merge(epochs_temp, by='EncounterKey', all.x=T) %>%
  merge(last_temp, by='EncounterKey', all.x=T) %>%
  merge(epochs_map, by='EncounterKey', all.x=T) %>%
  merge(last_map, by='EncounterKey', all.x=T) %>%
  merge(epochs_gcse, by='EncounterKey', all.x=T) %>%
  merge(last_gcse, by='EncounterKey', all.x=T) %>%
  merge(epochs_gcsv, by='EncounterKey', all.x=T) %>%
  merge(last_gcsv, by='EncounterKey', all.x=T) %>%
  merge(epochs_gcsm, by='EncounterKey', all.x=T) %>%
  merge(last_gcsm, by='EncounterKey', all.x=T) %>%
  merge(epochs_weight, by='EncounterKey', all.x=T) %>%
  merge(last_weight, by='EncounterKey', all.x=T) %>%
  merge(epochs_abx, by='EncounterKey', all.x=T) %>%
  merge(esofa %>% select('EncounterKey', contains('ESOFA')), by='EncounterKey', all.x=T) %>%
  merge(dx_trim, by='EncounterKey', all.x=T)

write.csv(output_df, 'sepsis_r01_epoch_data.csv', row.names=F)
```

```{r Evaluate missingness in epoch output}
prop_missing <- output_df %>%
  summarise_all(funs(mean(is.na(.)))) %>%
  gather(key, value)

prop_missing
```


```{r Pulse check frequency calculation}
pulse <- flowsheet_assess %>%
  filter(FlowsheetRowDisplayName == 'Pulse') %>%
  filter(EncounterKey %in% df_timestamp$EncounterKey) %>%
  merge(df_timestamp %>% select(EncounterKey, inclusion_timestamp), by='EncounterKey', all.x=T) %>%
  mutate(
    timestamp = ymd_hm(paste(DateKeyValue, TimeOfDayKeyValue)),
    pulse = as.numeric(Value)
  ) %>%
  filter(timestamp >= inclusion_timestamp & timestamp < inclusion_timestamp + hours(4)) %>%
  group_by(EncounterKey) %>%
  summarise(
    pulse_freq = n()/4
  )

pulse
```
```{r Save Environment}
save.image('sepsis_r01_environment.RData')
```

```{r}
# examine output_df's pulse columns, counting the number of non-NA values there are
# in epoch_1_pulse, epoch_2_pulse, epoch_3_pulse, epoch_4_pulse in a new column called pulse_count
output_df <- output_df %>%
  mutate(
    pulse_count = rowSums(!is.na(select(., epoch_1_pulse, epoch_2_pulse, epoch_3_pulse, epoch_4_pulse
                                        )))
  )

output_df %>% count(pulse_count)
```

